---
title: "P8105 Homework 1"
author: Heidi Lumish
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

#### Data frame creation
This code chunk creates a data frame comprised of a random sample of size 10 from a standard Normal distribution, a logical vector indicating whether elements of the sample are greater than 0, a character vector of length 10, and a factor vector of length 10 with 3 different factor levels:

```{r problem_1, message = FALSE}
library(tidyverse)

problem_1_df =
  tibble(
    norm_samp = rnorm(10, mean = 1),
    samp_pos = norm_samp > 0,
    color = c("red", "orange", "yellow", "green", "blue", "purple", "teal", "black", "brown", "white"),
    vec_factor = factor(c("low", "high", "med", "med", "low", "high", "low", "med", "high", "high"))
  )
```

The data frame created appears as follows:
```{r problem_1 data frame, echo = FALSE}
library(knitr)
kable(problem_1_df[1:10, ])
```

#### Mean of each variable
This code chunk takes the mean of each of the 4 variables in the data frame:

```{r means, collapse = TRUE}
mean_samp = mean(pull(problem_1_df, norm_samp))
mean_samp_pos = mean(pull(problem_1_df, samp_pos))
mean_color = mean(pull(problem_1_df, color))
mean_vec_fac = mean(pull(problem_1_df, vec_factor))
```
The code returns error messages for the variables "color" and "vec_factor," because it is not possible to take the mean of a character varaible ("color") or a factor variable ("vec_factor").

#### Convert to numeric variables
This code chunk attempts to convert the logical, character, and factor variables to numeric variables.

```{r, eval = FALSE}
problem_1_df$samp_pos <- as.numeric(problem_1_df$samp_pos)
problem_1_df$color <- as.numeric(problem_1_df$color)
problem_1_df$vec_factor <- as.numeric(problem_1_df$vec_factor)

```
The logical variable "samp_pos" and the factor variable "vec_factor" can be converted to numeric variables, however the character variable "color" cannot be converted to a numeric variable. This explains what happens when you try to take the mean, because while the logical variable was already stored as number and therefore a mean could be calculated, this was not the case for the character variable and the factor variable. While the factor variable can be converted to a numeric variable using the as.numeric function, character variables cannot be converted to numeric variables, and therefore a mean cannot be calculated.

## Problem 2

This code chunk imports the Penguins dataset.

```{r penguins}
data("penguins", package = "palmerpenguins")
```

#### Description of Penguins dataset

The data frame has `r nrow(penguins)` rows and `r ncol(penguins)` columns. The dataset includes the following variables: `r names(penguins)`.

The dataset was collected between `r min(pull(penguins, year))` and `r max(pull(penguins, year))`. There are a total of `r nrow(penguins)` penguins in the data set. There are `r sum(penguins$sex == "male", na.rm = TRUE)` male penguins and `r sum(penguins$sex == "female", na.rm = TRUE)` female penguins. Penguins came from three different islands (`r levels(pull(penguins, island))`) and included three different species (`r levels(pull(penguins, species))`).

The mean flipper length is `r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE, digits = 2), digits = 2)` mm, and the mean bill length is `r round(mean(pull(penguins, bill_length_mm), na.rm = TRUE, digits = 2), digits = 2)` mm.

#### Penguins scatterplot

This code chunk creates a scatterplot of flipper length (mm) vs. bill length (mm):

```{r scatterplot, collapse = TRUE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

```

The scatterplot shows a positive relationship between penguin flipper length and penguin bill length. The three penguin species are denoted by color.

This code chunk exports the scatterplot as a pdf file.
```{r scatterplot2, collapse = TRUE}
ggsave("penguins_scatterplot.pdf", height = 4, width = 6)
