---
title: "P8105 Homework 1"
author: Heidi Lumish
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

#### Data frame creation
This code chunk creates a data frame comprised of a random sample of size 10 from a standard Normal distribution, a logical vector indicating whether elements of the sample are greater than 0, a character vector of length 10, and a factor vector of length 10 with 3 different factor levels:

```{r problem_1, message = FALSE}
library(tidyverse)

problem_1_df =
  tibble(
    norm_samp = rnorm(10, mean = 1),
    samp_pos = norm_samp > 0,
    color = c("red", "orange", "yellow", "green", "blue", "purple", "teal", "black", "brown", "white"),
    vec_factor = factor(c("low", "high", "med", "med", "low", "high", "low", "med", "high", "high"))
  )
```

The data frame created appears as follows:
```{r problem_1 data frame, echo = FALSE}
library(knitr)
kable(problem_1_df[1:10, ])
```

#### Mean of each variable
This code chunk takes the mean of each of the 4 variables in the data frame:

```{r means, collapse = TRUE}
mean_samp = mean(pull(problem_1_df, norm_samp))
mean_samp_pos = mean(pull(problem_1_df, samp_pos))
mean_color = mean(pull(problem_1_df, color))
mean_vec_fac = mean(pull(problem_1_df, vec_factor))
```
The code returns error messages for the variables "color" and "vec_factor," because it is not possible to take the mean of a character varaible ("color") or a factor variable ("vec_factor").
